Synthesizing Symmetric Lenses

### 2019-02-16 papers/SymmetricLenses/main.tex
dunno

\newif\ifdraft\drafttrue  % set true to show comments
% \newif\ifdraft\draftfalse  % set true to show comments
\newif\ifanon\anonfalse    % set true to suppress names, etc.
\newif\ifappendices\appendicesfalse

\PassOptionsToPackage{usenames,dvipsnames,svgnames,table}{xcolor}
\documentclass[acmsmall,review,screen,anonymous]{acmart}
\settopmatter{printacmref=false,printccs=false}

@@

{\em Lenses} are programs that can be run both ``front to back'' and ``back
to front,'' allowing updates to both input and output data to be transferred
in both directions.
Since their introduction by~\citet{Focal2005-long}, lenses have been extensively
studied, extended, and applied.  Recent work~\cite{optician} has also demonstrated how
techniques from {\em type-directed program synthesis} can be used to
efficiently synthesize a simple class of lenses---so-called {\em
  bijective lenses} over string data---given a pair of types (regular
expressions) and a small number of examples.  \bcp{Why is the next line not
  indented?} \afm{I have truly no idea, deferring until I get a stackoverflow answer.}

We show how to extend this synthesis algorithm to a much wider class of
lenses, which we call {\em simple symmetric lenses}, including not only all
of the bijective lenses and the more widely used ``asymmetric''
lenses~\cite{Focal2005-long}, but also a rich subset of the full-blown
``symmetric lenses'' proposed by \citet{symmetric-lenses}.

Synthesizing simple symmetric lenses is substantially more challenging than
synthesizing bijective lenses: Since some of the information on each side
can be ``disconnected'' from the other side, there will in general be {\em
  many} lenses that agree with a given example. To guide the search process,
we use {\em stochastic regular expressions} and ideas from information
theory to estimate the amount of information propagated by a candidate lens,
in conjunction with user annotations identifying parts of the source and
target data structures as either {\em irrelevant} or {\em essential}.

Simple symmetric lenses are also of independent theoretical interest, being
the largest class of symmetric lenses that do not rely on persistent
internal state.

We describe an implementation of simple symmetric lenses and our synthesis
procedure as extensions to the Boomerang language~\cite{Boomerang}. We evaluate
its performance on 48 benchmark examples drawn from Flash Fill~\cite{flashfill},
Augeus~\cite{augeas2}, the bidirectional programming literature, and electronic
file format synchronization tasks. Our implementation can synthesize all these
lenses in under 30 seconds.
\end{abstract}

@@

\ifanon
%\authorinfo{}
%           {}
%           {}
\maketitle
% \vspace*{-6cm}
\else
\author{Anders Miltner}
\affiliation{
  \institution{Princeton University}
  \country{USA}
}
\email{amiltner@cs.princeton.edu}

@@

\maketitle
\fi

@@

\noindent
In today's data-rich world, similar information is often
stored in different places and in different formats. For instance, electronic
calendars come in iCalendar, vCalendar, and h-event
formats~\cite{calendar-formats}; US taxation data can be transmitted via Tax
XML (used by the US government for e-filing) or TXF (used by
TurboTax)~\cite{tax-formats}; and sewing machines can use embroidery files
in formats with names like PES, DST, and several
others~\cite{embroidery-formats}.
%
\bcp{I'm not satisfied with the embroidery example: It doesn't obviously
  correspond to a bidirectional situation (the sewing machine does not need
  to send back the embroidery file to anybody).  The tax example could be
  criticized on the same grounds.}
\afm{I still think that there are situations where the embroidery file makes
  sense (I'm sharing my design with a friend, who then tweaks it and sends it
  back. There may also need to file-specific information, where one contains
  parameters like how fast to embroider and some doesn't, to be honest I don't
  know much about embroidery though. Thinking about other examples, however.)}

One convenient way to maintain consistency between such varied data formats
is to use a \emph{lens}~\cite{Focal2005-long}---a bidirectional program that
can transform updates to data represented in a format $S$ to another format
$T$ and vice versa, providing ``round tripping'' guarantees that help ensure
information is not lost or corrupted as it is transformed back and forth
between different representations.
Domain-specific languages for writing lenses can facilitate building
principled data transformations.  For example, Boomerang~cite{boomerang} is a
language for expressing bidirectional string transformations.

\textit{Synthesizing} lenses can make the task even easier. Indeed, lens
languages, like Boomerang, are excellent targets for search-based synthesis, since their
specialized, sub-Turing-complete syntax and expressive type systems drastically
limit the space of possible programs.
%
In particular, the recent Optician~\cite{optician} enhancement to Boomerang has
demonstrated that it is feasible to synthesize quite complex examples, albeit
only for a restricted class of lenses---so-called bijective lenses.
Specifically, given source and target formats plus a small number of examples of
corresponding data, Optician's synthesis algorithm is guaranteed to find a
bijective lens matching the examples, if one exists. Indeed, it can often find
an appropriate lens in less than a second with no examples at all.
%, with just types describing the source and target formats.
% \dpw{I would cite the ICFP paper (cite string:
%   maina+:quotient-synthesis) as to appear but I'm not sure how to do that
%   without possibly violating the spirit of double-blind reviewing. ie, if it
%   hasn't appeared yet, how could we know the content if we aren't the authors?
%   Mentioning it in the 3rd person does not seem to suffice ...} \ksf{The ICFP
%   2018 website lists the paper; I think it is okay to cite it. The reviewer can
%   assume we contacted the authors for a copy. We could ask Stephanie for a
%   ruling if we want to be sure. One of the principles of double blind as adopted
%   by SIGPLAN is not to reduce the quality of the paper; omitting this citation
%   would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the
%   third person, even if that's a bit awkward, but otherwise citing it is OK.}
One reason that bijective lens synthesis can be so effective, even relative to
successful synthesis tools in other domains, is that there are typically not
very many bijections between a given pair of data formats. 
If the synthesis algorithm can
find any bijection at all, it is fairly likely to be the intended one.

However, the set of real-world use cases for bijective lenses is also
somewhat limited.  A more common situation is where two real-world data
formats share just {\em some} of their information content.  For instance,
two related database tables might share four out of five columns, but each
might have a different fifth column.  Or, one ad-hoc system configuration
file might include some format-specific meta data, such as a date or a
reference number, whereas the same configuration file on another operating
system does not.  Indeed among the 48 benchmark problems described in
Section \ref{sec:evaluation} below, all of the benchmarks taken from Flash
Fill~\cite{flashfill} and many of the benchmarks that synchronize between
two ad hoc file formats have this characteristic, 13 benchmarks in
total.

Can Optician's basic synthesis procedure be extended to a richer class of
lenses? In particular, what about \emph{symmetric
  lenses}~\cite{symmetric-lenses}---a much larger class that includes
bijective lenses, ``asymmetric'' lenses (where the transformation from a
source structure to a target can throw away information that must
then be restored when transferring from target to source), and even more flexibile
transformations that allow each side to retain information not present in
the other?

@@

The first is pragmatic. ``Classical'' symmetric lenses as presented
by ~\citet{symmetric-lenses} operate
over three structures: the ``left'' structure $X$, the ``right'' $Y$ and a complement
$C$ that contains the information not
present in either $X$ or $Y$. Logistically, these complements must be stored
and managed somehow.  More fundamentially, complements complicate
synthesis {\em specifications}---instead of just giving single examples of
source and target 
pairings, users would have to give longer ``interaction sequences'' to 
show how a lens should behave. To avoid these complexities, we define a
restricted variant 
of symmetric lenses, which we call \emph{simple symmetric lenses}. 
%only the usual source and target.
%
%, which we define in Section~\ref{sec:relationship}.
Intuitively, simple symmetric lenses are the complete set of symmetric
lenses that do not require external ``memory'' to recover data from past
instances of $X$ or $Y$ when making a round trip. They only need the most
recent instance.

@@

This leaves us with just the second difficulty with synthesizing symmetric
lenses: While the number of bijective lenses between two given formats is
typically tiny, the number of simple symmetric lenses is typically enormous.
If a na\"ive search algorithm just selects the first simple symmetric lens
it finds, this will generally \emph{not} be the one the user wanted.  We
need a new principle for identifying ``more likely'' lenses and a more
sophisticated synthesis algorithm that uses this principle to search the
space more intelligently. For these, we turn to information theory.

We consider a ``likely'' lens to be one that propagates ``a lot'' of information
from the left data format to the right and vice versa. Conversely, an unlikely
lens is one that requires a large amount of additional information to recover
either of the formats given the other. For many application scenarios, users
will want to synchronise formats with lots of shared information (and lenses
with little shared information are typically easier to write). Our tool is built
primarily for these complex scnearios, and so lenses that propagate more
information are preferable in our context to ones that propagate less.
%
This intuition is formalized using \emph{stochastic regular expressions}
(SREs)~\cite{stoch-rnn,stoch-def}, which simultaneously define a set of
strings and a probability distribution over those strings.  Using this
probability distribution, we can calculate the shared information content of
two formats and our measure of likelihood.

@@

With simple symmetric lenses and our SRE-based likelihood measure in hand, we
develop a new algorithm for synthesizing likely lenses. At its core, the
algorithm performs a type-directed search between the descriptions of the source
and target formats, measuring success using the likelihood measure. The
interesting complications arise from the need to deal with regular-expression
equivalences. There are infinitely many regular expressions equivalent to a
given one, and the lens returned by a type-directed search will in general
depend on which of the possible representations are chosen for its source and
target formats. Moreover, if Boomerang's type-equivalence rule is disallowed,
there may be {\em no} well-typed lens connecting the source and target formats
selected by the user: in general, the synthesis algorithm has to search through
equivalent regular expression types for which a suitable lens exists. To tame
this complexity, we divide the synthesis algorithm into two communicating search
procedures (Figure~\ref{fig:high-level-algorithm}), following~\citet{optician}.
The first, \Expand, uses rewriting rules to iteratively propose new pairs of
stochastic regular expressions equivalent to the original pair. The second,
\GreedySynth, takes a pair of SREs as input and uses a greedy, type-directed
algorithm to find a simple symmetric lens between them, returning the lens and
its likelihood score to \Expand. The whole synthesis algorithm heuristically
terminates when the most likely lens is no longer unlikely enough to justify
the time cost of searching through newly generated SREs.

@@

We have added this synthesis procedure to the Boomerang system and evaluated
its effectiveness on a benchmark suite of 48 problems, including data-cleaning
tasks taken from Flash Fill~\cite{gulwani-popl-2014}, view-maintenance tasks
taken from Augeas~\cite{augeas2}, and ad-hoc synchronization tasks from a
variety of other sources (mostly from the bidirectional programming literature).
The user can control the search process via examples and annotations
that informs the synthesis algorithm as to whether certain bits of information
are irrelevant, or must maintained.

In summary, our contributions are:
\begin{itemize}
%
\item We define {\em simple symmetric lenses}, a subclass of symmetric
lenses that do not maintain any hidden state (\S\ref{sec:overview}).  
%
We present a collection of combinators for building simple symmetric lenses and
describe how to calculate their likelihoods (\S\ref{sec:ssl}).
\item We give an algorithm for synthesizing lenses built from these combinators,
  using a novel application of stochastic regular expressions to guide the
  search process (\S\ref{sec:synthesis}) and prove that our metrics have deep
  semantic meaning.
%
\item We extend the Boomerang implementation with simple symmetric lenses,
  demonstrating that they integrate smoothly with other lens extensions such as
  quotient lenses~\cite{quotientlenses} and matching
  lenses~\cite{matchinglenses}. We evaluate our implementation, the effects of
  various optimizations, and our inclusion of relevance annotations on a set of
  48 lens synthesis benchmarks drawn from data cleaning, view maintenance, and
  file synchronization tasks. The system can synthesize simple symmetric lenses
  for all of the benchmarks in under 30 seconds (\S\ref{sec:evaluation}).
%
\item We show that simple symmetric lenses fall between asymmetric
lenses~\cite{Focal2005-long} and full symmetric
lenses~\cite{symmetric-lenses} in expressiveness.  Indeed, the
class of simple lenses can be characterized semantically as the subset of
full symmetric lenses for which a simple ``forgetfulness'' property holds
(\S\ref{sec:related}).
\end{itemize}
Along the
way, we show that {\em star-semiring equivalences} on regular expressions
remain valid when extended to stochastic regular expressions, preserving
probability distributions as well as languages (\S\ref{sec:sre}).
We close with related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).
Elided proofs can be found in the appendices\bcp{The appendices are pretty
  hideous and hard to navigate!!}
\ifappendices\else included with the supplemental material\fi.

@@

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{vlightyellow}}c|>{\columncolor{vlightgreen}}c|}
      \hline
      Management's data & HR's data \\
      \hline
      \listingA & \listingB \\
      \hline
      \cellcolor{lightyellow}Management's type & 
      \cellcolor{lightgreen} HR's type\\
      \hline
      \cellcolor{lightyellow}\listingC & \cellcolor{lightgreen}\listingD \\
      \hline
    \end{tabular}
  }
  \caption{Hypothetical example data files and corresponding regular expressions
  used by management and HR at a US company to represent employee salaries and health
  insurance providers, respectively. \bcp{The colors are a bit loud,
    especially since there is so little color elsewhere.}
  %
  \bcp{Suggest replacing ``Name,Company'' with ``First Last, Company'' (too
    many ``name''s flying around)}
  %
  \bcp{More substantially, thinking about this example makes me wonder about
    a whole different way of specifying simple symmetric lenses: rather than
    just giving corresponding source and target pairs, could we give triples
    of the form ``new source / old target / new target'' and ``new target /
    old source / new source''?  I know these would be a bit more work to
    write, but they would also give the synthesis procedure way more
    information to work with---it seems to me it should be able to do much
    less guessing about where things are disconnected.  Would this allow us
    to drop all the information theory?  Or would we still need it?  (In
    retrospect, this seems like a fairly obvious thing to try, so it seems
    important to try it and report on how it performs relative to what we
    propose here...)}\afm{Oh sorry, yes we can write such examples.  I personally found that typically the same information could be provided through a series of 2 creates (aka: if we want a disconnect: creater a = b; creater a' = b:  if we want an identity: creater a = a; creater a' = a').  However, we also permit putr and putl examples of the proposed form in the tool 
Much like other aspects of the work, any situation with examples ultimately only constrains the subcomponents of the lens corresponding to that specific parse tree (aka: if I give an example about A in A|B, it only constrains "A").  I don't think it removes the need for the information theory, as all the other non-constrained subcomponents of the lens can still be "anything."
}  }
  \label{fig:minimized-representations}
\end{figure}

@@

\subsection{Simple Symmetric Lenses}
Semantically, a simple symmetric lens between sets $X$ and
$Y$ comprises four functions subject to four
``round-tripping'' laws:

%\begin{enumerate}
%\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
%\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
%\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
%\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
%\end{enumerate}
\vskip 1.5ex
\begin{center}
\begin{minipage}[c]{.3\textwidth}
\begin{gather}
 \mbox{$\CreateR \OfType X \rightarrow Y$} \nonumber \\
 \mbox{$\CreateL \OfType Y \rightarrow X$} \nonumber \\
 \mbox{$\PutR \OfType X \rightarrow Y \rightarrow Y$} \nonumber \\
 \mbox{$\PutL \OfType Y \rightarrow X \rightarrow X$} \nonumber
\end{gather} 
\end{minipage} 
\hspace{.4in}
\begin{minipage}[c]{.5\textwidth}
\begin{align}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{align}
\end{minipage} 
\end{center}
\vskip 1ex
\noindent
The two \KW{create} functions are used to propagate shared information and fill
in default values when introducing new data (\EG, on \KW{create} the ``unk''
salary entry is added alongside the name to the management file when HR inserts
a new employee). The two \KW{put} functions propagate edits from one format to
the other by combining a new value from one with an old value from the other.
The notation $\Lens.\PutR$ extracts the \PutR function from simple symmetric
lens $\Lens$; similarly for \PutL. These four functions can be used to keep the
common information between two file types in synch. For example, if a new file
of the left-hand format is created, the \CreateR function will a new file in the
right-hand format. If the right-hand format is edited, the \PutL function can
update the left-hand format with the changed information.

The round-tripping laws guarantee that pushing an ``un-edited'' value from one
side (the result of a put or create) back through a lens in the other direction
will get back to exactly where we began. For example, consider,
\lstinline{emp_lens}, a lens between
\lstinline{emp_salaries} and
\lstinline{emp_insurance}. Applying
\lstinline{emp_lens}.\CreateR to a salary file creates a insurance file with the
same employee names in the file, with \lstinline{"UNK"} for each insurance
company. The round-tripping laws guarantee that using \lstinline{emp_lens}.\PutL
to put the generated insurance file into the original salary file will return
the original salary file, unmodified. If \KW{put}ting a string $s$ into a string
$t$ results in the original string $s$, then $s$ and $t$ are considered
\emph{synchronized}.

This simple symmetric lens definition differs from classical symmetric lenses in
that it does not involve a \emph{complement}. We give a detailed comparison in
\S\ref{sec:relationship}.

@@

To obtain SREs from the plain regular expressions that users write, we use a
heuristic that attempts to assign probability annotations that give each
string in the language equal probablity. Our synthesis algorithm could
equally well take as input a pair of stochastic regular expression rather than a
plain one, permitting users to specify a data-specific probability
distribution (either manually calculated or inferred from some dataset of
interest). However, we have found that our heuristic works well in practice,
so we have not explored these options.

>>>>

% \newif\ifdraft\drafttrue  % set true to show comments
\newif\ifdraft\draftfalse  % set true to show comments
\newif\ifanon\anonfalse    % set true to suppress names, etc.
\newif\ifappendices\appendicesfalse

\PassOptionsToPackage{usenames,dvipsnames,svgnames,table}{xcolor}
%\documentclass[acmsmall,review,screen,anonymous]{acmart}
\documentclass[acmsmall,review,screen]{acmart}
\settopmatter{printacmref=false,printccs=false}

@@

% BCP: To make sure we don't leave comments on!
\bcp{XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX }

{\em Lenses} are programs that can be run both ``front to back'' and ``back
to front,'' allowing data and updates to be transformed in two directions.
Since their 
introduction by~\citet{Focal2005-long}, lenses have been extensively
studied and applied.  Recent work~\cite{optician} has also demonstrated how
techniques from {\em type-directed program synthesis} can be used to
efficiently synthesize a very simple class of lenses---so-called {\em
  bijective lenses} over string data---given a pair of types (regular
expressions) and a small number of examples.

We show how to extend this synthesis algorithm to a much wider class of
lenses, which we call {\em simple symmetric lenses}, including both
bijective lenses and the more widely used ``asymmetric''
lenses~\cite{Focal2005-long}, as well as a rich subset of the full-blown
``symmetric lenses'' proposed by \citet{symmetric-lenses}.  Simple symmetric
lenses are of independent theoretical interest, being the largest class of
symmetric lenses that do not rely on persistent internal state.

Synthesizing simple symmetric lenses is substantially more challenging that
synthesizing bijective lenses: Since some of the information on each side
can be ``disconnected'' from the other side, there will in general be {\em
  many} lenses that agree with a given example.  To guide the search
process, we use {\em stochastic regular expressions} and ideas from
information theory to estimate the amount of information propagated by a
candidate lens.

We describe an implementation of simple symmetric lenses and our synthesis
procedure as extensions to the Boomerang language~\cite{Boomerang}. We evaluate
its performance on 48 benchmark examples drawn from Flash Fill~\cite{flashfill},
Augeus~\cite{augeas2}, the bidirectional programming literature, and electronic
file format synchronization tasks. After modest tuning, which is encouraged by
the interactive nature of the tool, our implementation can synthesize all these
lenses in under 30 seconds.
\end{abstract}

@@

%\ifanon
%\authorinfo{}
%           {}
%           {}
%\maketitle
% \vspace*{-6cm}
%\else
\author{Anders Miltner}
\affiliation{
  \institution{Princeton University}
  \country{USA}
}
\email{amiltner@cs.princeton.edu}

@@

\maketitle
%\fi

@@

\noindent
In today's big-data world, similar information is sometimes
stored in different places and in different formats. For instance, electronic
calendars come in iCalendar, vCalendar, and h-event
formats~\cite{calendar-formats}; US taxation data can be transmitted via Tax
XML (used by the US government for e-filing) or TXF (used by
TurboTax)~\cite{tax-formats}; and sewing machines can use embroidery files
in formats with names like PES, DST, and several
others~\cite{embroidery-formats}.
%
\bcp{I'm not satisfied with the embroidery example: It doesn't obviously
  correspond to a bidirectional situation (the sewing machine does not need
  to send back the embroidery file to anybody).  The tax example could be
  criticized on the same grounds.}
%
\bcp{Also, it would be nice to be able to give a quick example of a truly
  symmetric case---do any of these examples work for that?}\afm{h-event and
  vCalendar are truly symmetric. Even the fields related to location alone
  are truly symmetric (though it is subtle to see this).}

One convenient way to maintain consistency between these varied data formats
is to use a \emph{lens}~\cite{Focal2005-long}---a bidirectional program that
can transform updates to data represented in a format $S$ to another format
$T$ and vice versa, providing ``round tripping'' guarantees that help ensure
information is not lost or corrupted as it is transformed back and forth
between different representations.

While domain-specific languages for writing lenses can facilitate building
principled data transformations, \textit{synthesizing} lenses 
can make the task even easier.  Indeed, lens languages offer excellent
targets for synthesis: their specialized, sub-Turing-complete syntax
and expressive type systems drastically
reduce the space of possible programs, making search-based synthesis easier. 
%
Recent work~\cite{optician} has exploited this observation, demonstrating that
it is feasible to synthesize quite complex examples, albeit only for a
restricted class of lenses---so-called bijective lenses. Specifically, given 
source and target formats plus a small number of
examples of corresponding data, the bijective synthesis algorithm will find a
bijective lens matching the examples, if one exists. Often,
it can find an appropriate lens in less than a second with
no examples at all.
%, with just types describing the source and target formats.
% \dpw{I would cite the ICFP paper (cite string:
%   maina+:quotient-synthesis) as to appear but I'm not sure how to do that
%   without possibly violating the spirit of double-blind reviewing. ie, if it
%   hasn't appeared yet, how could we know the content if we aren't the authors?
%   Mentioning it in the 3rd person does not seem to suffice ...} \ksf{The ICFP
%   2018 website lists the paper; I think it is okay to cite it. The reviewer can
%   assume we contacted the authors for a copy. We could ask Stephanie for a
%   ruling if we want to be sure. One of the principles of double blind as adopted
%   by SIGPLAN is not to reduce the quality of the paper; omitting this citation
%   would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the
%   third person, even if that's a bit awkward, but otherwise citing it is OK.}
One reason that bijective lens synthesis is so effective, even relative to
successful synthesis tools in other domains, is that there are typically not
very many bijections between related data formats. 
Hence, if the synthesis algorithm can
find any bijection at all, it is likely to be the right one.

However, many real-world data formats share just part of their information
content with other representations.  For instance, two related
database tables might share four out of five columns, but each might have a
different fifth column. 
Or, one ad-hoc system configuration file might include some format-specific meta
data, such as a date or a reference number, whereas the same configuration
file on another operating system does not. 
Indeed of the 48 benchmark problems described in Section \ref{sec:evaluation},
all of the benchmarks taken from Flash Fill~\cite{flashfill} and many of the
benchmarks that synchronize between two ad hoc file formats have this
characteristic, 13 benchmarks in total. Consequently, synthesizing only strict
bijections is often not useful in practice.

Could we extend synthesis tools to more kinds of lenses?  In particular,
what about \emph{symmetric lenses}~\cite{symmetric-lenses}, a class that
includes bijective lenses, ``asymmetric'' lenses where the transformation
from $S$ to $T$ can throw away information that must then be restored when
returning from $T$ to $S$, and even more flexibile transformations that
allow each side to retain information not present in the other?

\bcp{We use $S$ and $T$ here but $X$ and $Y$ below.}

@@

The first is pragmatic. ``Classical'' symmetric lenses as presented
by ~\citet{symmetric-lenses} operate
over three structures: the source $S$, the target $T$ and a complement
$C$ that contains the information not
present in either $S$ or $T$. Logistically, these complements must be stored
and managed somehow.  More fundamentially, complements complicate
synthesis {\em specifications}---instead of just giving single examples of
source and target 
pairings, users would have to give longer ``interaction sequences'' to 
show how a lens should behave. To avoid these complexities, we define a
restricted variant 
of symmetric lenses, which we call \emph{simple symmetric lenses}. 
%only the usual source and target.
%
%, which we define in Section~\ref{sec:relationship}.
Intuitively, simple symmetric lenses are the complete set of symmetric
lenses that do not require external ``memory'' to recover data from past
instances of $S$ or $T$ when making a round trip. They only need the most
recent instance.

@@

However, we are still left with a second difficulty with synthesizing
symmetric lenses: While the number of bijective lenses between two related
formats is typically tiny, the number of simple symmetric lenses is
typically huge.  Thus, if a na\"ive search algorithm selects the first
simple symmetric lens it finds, this will generally \emph{not} be the one we
want.  We need a new principle for identifying ``more likely'' lenses and a
more sophisticated synthesis algorithm that uses this principles to search
the space more intelligently.

For these, we turn to information theory. 
We consider a ``likely'' lens to be one that propagates ``a lot'' of
information from source to 
target and vice versa. Conversely, an unlikely lens requires a
large amount of additional information to recover the target given a
source.  
This categorization follows from the intuition
that users will typically want to synchronise formats with lots
of shared information, and so lenses that propagate
more information are preferable to ones that propagate less.
%
This intuition is formalized using \emph{stochastic regular expressions}
(SREs)~\cite{stoch-rnn,stoch-def}, which simultaneously define a set of
strings and a probability distribution over those strings.  Using this
probability distribution, we can calculate the shared information content of
two formats\bcp{It's not obvious at this point what the {\em shared}
  information content has to do with anything.} and our measure of likelihood.

@@

With simple symmetric lenses and our SRE-based likelihood measure in
hand, we develop a new algorithm for synthesizing likely lenses.
At its core, the algorithm performs a type-directed search between the 
descriptions of the source and target formats, measuring success using
the likelihood measure.
The interesting complications arise from the need to deal with
regular-expression equivalences.  There are infinitely many regular
expressions equivalent to a given one, and the lens returned by a
type-directed search will depend which of the infinitely many possible
representations are chosen for its source and target formats.  Moreover,
there may be {\em no} lens connecting the 
source and target formats selected by the user: in general, the synthesis
algorithm has to search to find equivalent ones for which a suitable lens
exists.  
To tame this complexity, we divide the synthesis algorithm into two
communicating search procedures (Figure~\ref{fig:high-level-algorithm}),
following~\citet{optician}. 
The first, \Expand, uses rewriting rules to iteratively
propose new pairs of stochastic regular expressions equivalent to the
original pair.
The second, \GreedySynth, takes a pair of SREs as input
and uses a greedy, type-directed algorithm to find a simple symmetric
lens between them, returning the lens and its likelihood score
to \Expand.  The whole synthesis algorithm terminates when comparing the
likilihood score of the best lens found so far with 
the complexity of newly generated SREs 
suggests that further searching is unlikely to be fruitful.

@@

We added this synthesis procedure to the Boomerang implementation and
evaluated its effectiveness on a benchmark suite of 48 problems, including
data-cleaning tasks taken from Flash Fill~\cite{gulwani-popl-2014},
view-maintenance tasks taken from Augeas~\cite{augeas2}, and ad-hoc
synchronization tasks from a variety of other sources (mostly from the
bidirectional programming literature).  The ``programmer'' can control the
search process via examples and a {\em tuning parameter} that governs how
much effort is spent trying to improve on existing solutions.

In summary, our contributions are:\bcp{The section references do not proceed
  in order, so this doesn't work very well as an outline of the paper.
  Should we try to reorder the contributions?  Write a separate outline?
  Live with it as-is?}\afm{Personally, I don't see it as a problem.  It's
  nice when the contributions mirror the paper, I am not jarred when they
  don't.}\bcp{But then the reader has no birds-eye view of the paper at
  this point.  I tried reorganizing, which I think also helps clarify the
  contributions.} 
\begin{itemize}
%
\item We define {\em simple symmetric lenses}, a subclass of symmetric
lenses that do not maintain any hidden state (\S\ref{sec:overview}).  
%
We
present a collection of combinators for building simple symmetric lenses and
describe how to calculate their likelihoods (\S\ref{sec:ssl}).
\item We give
an algorithm for synthesizing lenses built from these combinators, using a 
novel application of stochastic regular expressions to guide the search
process (\S\ref{sec:synthesis})\bcp{``and prove it correct''???}.  
%
\item We extend the Boomerang implementation with simple symmetric lenses,
demonstrating that they integrate smoothly with other lens extensions such
as quotient lenses~\cite{quotientlenses} and matching
lenses~\cite{matchinglenses}.  We evaluate our implementation and the
effects of various optimizations and the tuning parameter on a set of 48
lens synthesis benchmarks drawn from data cleaning, view maintenance, and
file synchronization tasks.  With only modest interactive tuning, the system
can synthesize simple symmetric lenses for all of the benchmarks in under 30
seconds (\S\ref{sec:evaluation}).
%
\item We show that simple symmetric lenses fall between asymmetric
lenses~\cite{Focal2005-long} and full symmetric
lenses~\cite{symmetric-lenses} in expressiveness.  Indeed, the
class of simple lenses can be characterized semantically as the subset of
full symmetric lenses for which a simple ``forgetfulness'' property holds
(\S\ref{sec:related}).
\end{itemize}
Along the
way, we show that {\em star-semiring equivalences} on regular expressions
remain valid when applied to stochastic regular expressions, preserving
probability distributions as well as languages (\S\ref{sec:sre}).
We close with related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).
Elided proofs can be found in the appendices\bcp{The appendices are pretty
  hideous and hard to navigate!!}
\ifappendices\else included with the supplemental material\fi.

@@

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{vlightyellow}}c|>{\columncolor{vlightgreen}}c|}
      \hline
      Management's data & HR's data \\
      \hline
      \listingA & \listingB \\
      \hline
      \cellcolor{lightyellow}Management's type & 
      \cellcolor{lightgreen} HR's type\\
      \hline
      \cellcolor{lightyellow}\listingC & \cellcolor{lightgreen}\listingD \\
      \hline
    \end{tabular}
  }
  \caption{Hypothetical example data files and corresponding regular expressions
  used by management and HR at a US company to represent employee salaries and health
  insurance providers, respectively. \bcp{The colors are a bit loud,
    especially since there is so little color elsewhere.}
  %
  \bcp{Suggest replacing ``Name,Company'' with ``First Last, Company'' (too
    many ``name''s flying around)}
  %
  \bcp{More substantially, thinking about this example makes me wonder about
    a whole different way of specifying simple symmetric lenses: rather than
    just giving corresponding source and target pairs, could we give triples
    of the form ``new source / old target / new target'' and ``new target /
    old source / new source''?  I know these would be a bit more work to
    write, but they would also give the synthesis procedure way more
    information to work with---it seems to me it should be able to do much
    less guessing about where things are disconnected.  Would this allow us
    to drop all the information theory?  Or would we still need it?  (In
    retrospect, this seems like a fairly obvious thing to try, so it seems
    important to try it and report on how it performs relative to what we
    propose here...)}  }
  \label{fig:minimized-representations}
\end{figure}

@@

\subsection{Simple Symmetric Lenses}
Semantically, a simple symmetric lens between sets $X$ and
$Y$ comprises four functions that collectively satisfy four
``round-tripping'' laws:

%\begin{enumerate}
%\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
%\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
%\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
%\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
%\end{enumerate}
\vskip 1.5ex
\begin{minipage}[c]{.3\textwidth}
\begin{gather}
 \tag{1}
 \mbox{$\CreateR \OfType X \rightarrow Y$}\\
 \tag{2}
 \mbox{$\CreateL \OfType Y \rightarrow X$}\\
 \tag{3}
 \mbox{$\PutR \OfType X \rightarrow Y \rightarrow Y$}\\
 \tag{4}
 \mbox{$\PutL \OfType Y \rightarrow X \rightarrow X$}
\end{gather} 
\end{minipage} 
\begin{minipage}[c]{.6\textwidth}
\begin{align}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{align}
\end{minipage} 
\vskip 1ex
\bcp{I want less space between the put/create laws and their names.}
\noindent
The two \KW{create} functions are used to propagate shared information and
fill in default values when introducing new data (\EG, the ``unk'' salary
entry is added alongside the name to the management file when HR inserts a
new employee). The two \KW{put} functions propagate edits from one format to
the other by combining a new value from one with an old value from the
other.  The round-tripping laws guarantee that pushing an ``un-edited''
value from one side (the result of a put or create) back through a lens in
the other direction will get back to exactly where we began.

This definition differs from classic symmetric lenses in that it does not
involve a \emph{complement}.  We give a detailed comparison in
\S\ref{sec:relationship}.

@@

To obtain SREs from the plain regular expressions that we expect users to
supply, we use a heuristic that attempts to assign probability annotations that
give each string in the language equal probablity. Our algorithm could also take
as input a stochastic regular expression rather than a plain one,
which would allow
the user to specify a data-specific probability distribution, either manually
calculated or inferred from a large dataset. We found our heuristic
works well in practice, so we have not explored these options.
